---
title: "p8105_hw1_zw2541"
author: "Zixu_Wang"
date: "9/15/2018"
output: github_document
---

# Problem 1

## Create a data frame
This data frame comprised of:

*  A random sample of size 10 from a uniform[0, 5] distribution
*  A logical vector indicating whether elements of the sample are greater than 2
*  A (length-10) character vector
*  A (length-10) factor vector

```{r, message = FALSE}
library(tidyverse)
```

```{r p1_dataframe}
set.seed(1)
p1_df = tibble(
  vec_numeric = runif(10, min = 0, max = 5),
  vec_logical = vec_numeric > 2,
  vec_char = c("zhao", "qian", "sun", "li", "zhou", "wu", "zheng", "wang", "feng", "chen"),
  vec_factor = factor(c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"))
)

class(p1_df$vec_numeric)
class(p1_df$vec_logical)
class(p1_df$vec_char)
class(p1_df$vec_factor)
```

## Take the mean of each variable in the dataframe

```{r mean_of_each_variable}
mean(p1_df$vec_numeric)
mean(p1_df$vec_logical)
mean(p1_df$vec_char)
mean(p1_df$vec_factor)
```
Result: I successfully calculate the mean of numeric and logical variables, but I can't get the mean of character and factor variables.

Explanation: Since the numeric variable is a set of numbers, it is quite easy to calculate their mean. For the logical variable, the system recognize TRUE as 1 and FALSE as 0, so it can calculate the mean of a set of number 1 and 0. However, the system can't calculate the mean of character and factor variables, because these two variables are a set of words instead of numbers.

## Convert variables from one type to another

The code chunk showed below will convert logical, character, and factor variables to numeric variable:
```{r convert_variables_type, warning = FALSE, results = FALSE}
as.numeric(p1_df$vec_logical)
as.numeric(p1_df$vec_char)
as.numeric(p1_df$vec_factor)
```
For the logical variable, the as.numeric function converts TRUE to numeric value 1 and FALSE to numeric value 0. And this function can't convert character variable to numeric variable. And as.numeric function converts favtor variable to different value of numbers from 1 to 10.


These two code chunks showed below convert the character variable from character to factor to numeric, and similarly convert the factor variable from factor to character to numeric. 
```{r double_convert_variable_type}
as.numeric(as.factor(p1_df$vec_char))
as.numeric(as.character(p1_df$vec_factor))
```
The first code chunk successfully converts the character variable to factor to numeric. However, the second chunk still can't convert the character variable to numeric.


# Problem 2

## Create a data frame
This data frame comprised of:

*  x: a random sample of size 1000 from a standard Normal distribution
*  y: a random sample of size 1000 from a standard Normal distribution
*  A logical vector indicating whether the x + y > 0
*  A numeric vector created by coercing the above logical vector
*  A factor vector created by coercing the above logical vector

```{r p2_dataframe}
p2_df = tibble(
x = rnorm(1000),
y = rnorm(1000),
vec_logical = x + y > 0,
vec_numeric = as.numeric(vec_logical),
vec_factor = as.factor(vec_logical)
)
```

## Inline R code
As the data frame showed above, *the size of the dataset* is `r nrow(p2_df)`, *the mean of x is* `r mean(p2_df$x)`, and *the median of x* is `r median(p2_df$x)`. Besides, *the proportion of cases for which the logical vector is TRUE* is `r sum(p2_df$vec_logical) / length(p2_df$vec_logical)`. (Explanation: When I calculating the sum of logical variable, the system may recognize the TRUE as numeric value 1 and the FALSE as numeric value 0. So, the sum of logical variable means the number of the cases that the logical vector is TRUE. And by making this sum devided by the length of this logical variable which is 1000, we can get the proportion of cases for which the logical vector is TRUE.)

## Scatterplot of y vs x

### scatterplot of y vs x with the color points using the logical variable
```{r Scatterplot_logical}
ggplot(p2_df, aes(x = x, y = y, color = p2_df$vec_logical)) + geom_point()
ggsave("scatterplot_with_color_points_of_logical_variable.png")
```
From the scatterplot above, we could find that nearly 50% of the logical variables are true and 50% are false.

### scatterplot of y vs x with the color points using the numeric variable
```{r Scatterplot_numeric}
ggplot(p2_df, aes(x = x, y = y, color = p2_df$vec_numeric)) + geom_point()
```
From the scatterplot above, we could find that nearly 50% of the numeric variables equal 1 and 50% equal 0.

### scatterplot of y vs x with the color points using the factor variable
```{r Scatterplot_factor}
ggplot(p2_df, aes(x = x, y = y, color = p2_df$vec_factor)) + geom_point()
```
From the scatterplot above, we could find that nearly 50% of the factor variables are ture and 50% are false.
