---
title: "p8105_hw1_zw2541"
author: "Zixu_Wang"
date: "9/15/2018"
output: github_document
---

# Problem 1

## Create a data frame
This data frame comprised of:

*  A random sample of size 10 from a uniform[0, 5] distribution
*  A logical vector indicating whether elements of the sample are greater than 2
*  A (length-10) character vector
*  A (length-10) factor vector

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
```

```{r p1_dataframe}
set.seed(1)
p1_df = tibble(
  vec_numeric = runif(10, min = 0, max = 5),
  vec_logical = c(vec_numeric > 2),
  vec_char = c("zhao", "qian", "sun", "li", "zhou", "wu", "zheng", "wang", "feng", "chen"),
  vec_factor = factor(c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"))
)
```

## take the mean of each variable in the dataframe

```{r mean_of_each_variable}
mean(p1_df$vec_numeric)
mean(p1_df$vec_logical)
mean(p1_df$vec_char)
mean(p1_df$vec_factor)
```
Result: I successfully calculate the mean of numeric and logical variables, but I can't get the mean of character and factor variables.

Explanation: Since the numeric variable is a set of numbers, it is quite easy to calculate their mean. For the logical variable, the system recognize TRUE as 1 and FALSE as 0, so it can calculate the mean of a set of number 1 and 0. However, the system can't calculate the mean of character and factor variables, because these two variables are a set of words instead of numbers.

##convert variables from one type to another

The code chunk showed below will convert logical, character, and factor variables to numeric variable:
```{r convert_variables_type, warning = FALSE, results = FALSE}
as.numeric(p1_df$vec_logical)
as.numeric(p1_df$vec_char)
as.numeric(p1_df$vec_factor)
```
For the logical variable, the as.numeric function converts TRUE to numeric value 1 and FALSE to numeric value 0. And this function can't convert character variable to numeric variable. And as.numeric function converts favtor variable to different value of numbers from 1 to 10.


These two code chunks showed below convert the character variable from character to factor to numeric, and similarly convert the factor variable from factor to character to numeric. 
```{r double_convert_variable_type}
as.numeric(as.factor(p1_df$vec_char))
as.numeric(as.character(p1_df$vec_factor))
```
The first code chunk successfully converts the character variable to factor to numeric. However, the second chunk still can't convert the character variable to numeric.


 
